|

### 2021-10-06 03-Chp-Systematic-Review.Rmd
Working version

# Systematic review of all evidence available on the association between blood lipids and dementia outcomes {#sys-rev-heading}

@@

For

A random-effects meta-analysis model was employed to combine the different included studies. <!-- TODO Need more information on the difference between the two approaches, and their formula here, and description of different methods and quarrels regarding distinction fixed effect vs random effects meta-analyses. --> The fixed-effect method was implemented as:

<!-- TODO Check these against those in Ben's thesis paper -->

Results were stratified into subgroups on the basis of the overall risk of bias assessment, and summary estimates for each subgroup, in addition to an overall effect estimate, are displayed. Additional descriptive statistics are presented, while prediction intervals are shown as a dotted line around the overall effect estimate. Finally, a test subgroup differences between studies at different levels of risk of bias was performed (see subsequent Section \@ref(sys-rev-visualising-results)).

Random-effects analyses assume that differences in study design and execution mean that the study effects are normally distributed around a common treatment effect ($\mu$) with a variance of $\tau^2$

\begin{equation}
  \theta_i \sim \mathcal{N}(\mu, \tau^2) (\#eq:eq9)
\end{equation}

<!-- Random-effects methods estimate $\tau^2$ to estimate the common effect ($\mu$). There are many estimators of $\tau^2$ with varying assumptions [@Veroniki2016; @Bakbergenuly2020], the details of which are beyond the scope of this article. The estimate of $\tau^2$ is incorporated into the weighting of each study -->

\begin{equation}
  w_i = \frac{1}{v_i + \hat{\tau^2}} (\#eq:eq10)
\end{equation}

<!-- A random-effects meta-analysis allows for study-specific effects, and assumes that the true effect estimated by each study varies and is assumed to be randomly and normally distributed on the transformed scale (e.g. log-ORs) between studies, with a variance of 𝜏2. The between-study variance is estimated (usually using the “method of moments” or DerSimonian and Laird estimate), and used to modify the weights of studies when calculating the summary effect estimate and variance. The weight of each study is 1𝑣𝑖+𝜏2 where 𝑣𝑖 is the variance of the 𝑖𝑡ℎ study and 𝜏2 is the estimate of the between-study variance. So long as the between-study variance is larger than 0, smaller studies will be given more weight in random-effects meta-analysis than in a fixed-effect meta-analysis. If the between-study variance is 0, then the random-effects model is equivalent to the fixed-effect model. -->
<!-- In both fixed-effect and random-effects inverse-variance meta-analysis, the summary effect estimate and variance are estimated in the same way. The effect estimates (𝜃𝑖) from each study can be (transformed) ORs, relative risks (RRs), standardised mean differences (SMDs) or other types of estimates: -->
<!-- 𝜃̂𝑆𝑢𝑚𝑚𝑎𝑟𝑦=⁡Σ𝑤𝑖𝜃𝑖Σ𝑤𝑖 -->
<!-- (2) -->
<!-- 𝑉𝑎𝑟𝑖𝑎𝑛𝑐𝑒⁡𝑜𝑓⁡𝜃̂𝑆𝑢𝑚𝑚𝑎𝑟𝑦=⁡1Σ𝑤𝑖 -->
<!-- (3) -->
<!-- where 𝜃𝑖 is effect estimate and 𝑤𝑖 is the weight from the 𝑖𝑡ℎ study on an approximately normal scale. -->

<!-- \begin{equation} -->
<!--   \theta_i = \mu + u_i -->
<!--   (\#eq:meta-analysis-random) -->
<!-- \end{equation} -->

<!-- \begin{equation} -->
<!--   $weighted\ average = \frac{\sum Y_i (1/SE_i^2)}{\sum(1/SE_i^2)}$ -->
<!--   (\#eq:meta-analysis-fixed) -->
<!-- \end{equation} -->

<!-- where ... -->

<!-- TODO Need equation for random effects too. See Sean's thesis here. -->

@@

<!-- Will also need description of dose-response meta-analysis here, with reference to forthcoming book Julian sent on. -->

@@

Following best practice, studies judged to be at critical risk of bias should be excluded from any quantitative analyses. <!-- TODO CITATION NEEDED --> `r num_to_words(n_critical,T)` observational studies were excluded on this basis, predominantly due to a lack of adjustment for any confounders (i.e. the study reported unadjusted estimates).`r critical_citations`

@@

`r num_to_words(obsHyperchol$VaD$n,T)` studies reported on the association of hypercholesterolemia with vascular dementia and provided weak evidence for an effect (`r obsHyperchol$VaD$estimate`, Figure \@ref(fig:obsHyperVaD))

@@

<!-- TODO Cite excluded studies and list exclusion reasons -->

@@

Several studies were excluded from the dose-response meta-analysis, as the number of cases/controls per dose group could not be calculated and the corresponding author for the study did not respond to clarification requests. The results from the dose response analysis can be seen in Figure \@ref(fig:lipidsDoseResponse).

@@

Data on another potentially important source of heterogeneity
Plus other such as education level and baseline cognitive scores, but data were either not reported for most studies or when reported, were too diverse to synthesize.

@@

<!----------------------------------------------------------------------------->
(ref:funnelStatinsAny-cap) Funnel plot of results examining the relationship between statins and any dementia

(ref:funnelStatinsAny-scap) Funnel plot of results examining the relationship between statins and any dementia

```{r funnelStatinsAny, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:funnelStatinsAny-cap)', out.width='100%', fig.scap='(ref:funnelStatinsAny-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_rct_statins_Dementia.png"))
```

@@

The risk of bias due to missing evidence in each synthesis is shown beside the overall summary diamond in each forest plot presented above. For randomised controlled trials and non-randomised studies of interventions, the risk of missing. However, there was substantial evidence that results were selectively reported in studies examining the effect of lipid fractions on dementia outcomes (as shown in Figures \@ref(fig:lipidFractionsDementia), \@(lipidFractionsAD) and lipidFractionsAD). A key example is

@@

As show in Figure \@ref(prisma-flow-fig), the number of hits returned by the preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). From these hits, three preprinted reports of eligible studies were included in the review, of which two described unique studies not captured by the main search.[@andrews2019; @so2017]

All identified preprinted studies were. However, including preprints did provide useful additional evidence value in a number of meta-analysis. For example, re-analysing the results of available for the effect of HMGCR SNPS on Alzheimer's disease (Figure \@ref) in Mendelian randomisation studies, illustrates that the result only available from a preprint contributed `r mrStatinPreprintWeight` of the weight in a fixed-effect meta-analysis.

@@

<!----------------------------------------------------------------------------->
(ref:preprintWeights-caption) Amount of information added by inclusion of preprinted results, estimated from the weight assigned to the study in a random-effect meta-analysis

(ref:preprintWeights-scaption) Added information due to inclusion of preprints

```{r preprintWeights-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

Investigation of the publication status of the two unique preprints indicated that one has since been published (in late 2019 date),[@andrews2019; @andrews2021]. The final preprint has not yet been published.[@so2017]

@@

<!-- TODO Check inclusion of Rockwood 2002:  -->
This review has presented a summary of the available evidence on the association between lipids, and treatments that modify lipids such as statins, and the subsequent risk of dementia.

The discussion seeks to  to summarise the key findings in terms of literature sources and results as reported. A detailed comparison across the evidence sources, exposure measures and sources of bias reported here is presented as part of the triangulation exercise (Chapter \@ref(tri-heading)).

@@

There was some evidence of protective effect of statins on all-cause and Alzheimer's disease dementia when looking at solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings may be a result of heterogenity in exposure (e.g. mid-life in studies of lipids with late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

Some evidence that age has impact on observed lipid-dementia relationship - use to link

The majority of studies were non-randomised studies of lipids, or treatments that affect lipid levels such as statins. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the long follow-up, necessary due to the long latent period of the condition, makes trials logistically challenging and financial expensive. Similarly, Mendelian randomisation is a comparatively new study design (as illustrated in Figure \@ref(fig:typeByYear)), and so only appears in the literature in recent years, driven by the availability of Alzheimer's disease summary genome wide association studies (GWAS) that form the basis of two-sample Mendelian randomisation approach.

@@

A common theme across the evidence base was a lack of data on the association of vascular dementia. This is particularly interesting given that lipids and statins are primarily related to vascular disease. __Can't look forward to other chapter__ There is the potential that studies encountered similar difficulties in address the unexpected results observed in the CPRD analysis in Chapter \@ref(cprd-analysis-heading), likely due to confounding by indication, and so may suffer from the "file-drawer effect".[@rosenthal1979] For vascular dementia, few Mendelian randomisation studies examined this outcome, primarily because of the absence (until recently) of GWAS of this outcome.

<!--- 
QUESTION
This does introduce some conflicts with other studies where a change score has been dichotomised in order to have a binary dementia or not dementia outcome. - No these studies should not be included, as not recognised criteria.
-->

<!-- __Exclusion of the PROSPER trial__ Will need to have a good bit here on the exclusion of this RCT, particularly as it is included in the Cochrane review on the topic. Cross-reference with the meta-bias section in the early section, on the impact of only included studies that sought to make a definitive diagnosis, rather than using change in cognitive assessment scales. -->

Of note, this review did not include the commonly cited PROSPER study, which examined the effect of pravastatin on CVD risk,[@shepherd2002a] reporting on cognitive outcomes as one of several secondary outcomes. While widely cited in relation to the effect of statins on dementia and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial only reported on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up. Though an useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using recognised criteria, as cognitive tests should feed into a broader diagnostic pathway (see Section \@ref(diagnostic-criteria)). As such, this trial did not met the inclusion criteria for this review.

@@

Questions over missing results - evidence from one of the conference abstract analysis pairs that a non-significant results are being suppressed.[@yamada2009; @yamada2009a] In addition, there were some concerns over the potential for estimates to be missing from the meta-analysis of observational studies not at random, given the preferential reporting of significant results observed in a number of analysis (see Section \@ref(risk-of-bias-res)).

Generalisability was aided by the large number of studies performed using the Taiwan health database, which represents a good sources of evidence on non-Caucasian populations

@@

__This section will be completed once Georgia has completed her analysis, and can also be cross-references with the meta-meta analysis published is Brain Sciences recently.__

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018a; @kuzma2018a] However, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies, and explore the heterogeneity of results across different levels of risk of bias levels. Some previous reviews did assess risk of bias, but used non-domain based assessment tools, such as the Newcastle-Ottowa scale.[@anstey2015; @poly2020]

Similar the on previous review of Mendelian randomisation studies examining risk factors for Alzhiemer's disease was conducted prior to the majority of MR studies included in this review being published, and extracted results including SNPs in APoE4 (see the following section for a discussion of the bias this introduces).

However, despite these differences in timescales and methodology, the duplication of work across reviews (including this review) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch could have been employed. Known as an umbrella review, or review-of-reviews, these studies use other systematic reviews rather than primary studies as the unit of analysis.[@aromataris2015; @smith2011] This approach would have enabled more efficient identification of relevant primary studies to which the methods which sets this review apart from other published reviews could have been applied.

@@

One of the strengths of this review is it's inclusion of Mendelian randomisation studies as a source of evidence.

Mendelian randomisation is a powerful analytical technique, using natural variation in participants genomes to (assuming the assumptions of the method are valid), though it's inclusion as an acceptable study design in this review was complicated by a number of factors.

@@

```{r mrAssumptions, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:mrAssumptions-cap)', out.width='100%', fig.scap='(ref:mrAssumptions-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/mrAssumptions.png"))
```
<!----------------------------------------------------------------------------->

Firstly, this study design is relatively new, particularly when compared to randomised trials or cohort studies. Figure \@ref(fig:typeByYear) demonstrates that Mendelian randomisation studies only begin to appear in the evidence base much later than NRSE/NRSI. As such, the process and tools for systematically assessing them are not as well developed, likely due to the limited availabilt of large scale GWAS datasets needed for two sample MR. A key example of this is in the absence of validated search filters for Mendelian randomisations studies. This limitation is further complicated by the varying terminology used to describe the method, particularly in the early years of it's application.

Additionally, there is currently no widely used risk-of-bias assessment tool for Mendelian randomisation studies. A recent commentary provided a checklist  interpreting Mendelian randomisation studies, this guide includes reporting items in their quality checklist. While reporting quality is important, it is a separate consideration to internal validity, as discussed in Section \@ref(). Similarly, a previous review of Mendelian randomisation studies used the Q-Genie tool, which was validated to assess the quality of genetic association studies in meta-analysis.[@sohani2015] While this tool addresses the studies used, it does not access the additional methodological considerations of the analysis of the Mendelian randomisation analysis itself. <!-- TODO Read this tool. --> For this review, I utilised the best available author-devised tool, sourced on a recent review of systematic reviews of Mendelian randomisation studies.

<!-- TODO Need to add some text to the ROB section of this chapter on distinction between reporting and quality -->

<!-- Problems with overlapping samples in reviews of Mendelian randomisation studies in that may be double counting participants -->
<!-- TODO Check discussion on this point from Matt Lee thesis, in email -->
<!-- TODO Email George with query on dual counting, following review of email with Julian -->

@@

Useful future work to improve the methodology for inclusion of Mendelian randomisation studies in systematic reviews should involve the development of a validated search filter for this study design.[@waffenschmidt2020; @wagner2020] Alternatively, in better-resourced reviews, a dedicated search for "risk factors" and "dementia" and "Mendelian randomisation", followed by manual review of studies that look across multiple risk factors, would be advisable. This was not feasible in the context of this review, given the large number of records to be screened, even when using study design filters (n=`r comma(16109)`). Additionally, the value of methods that supporting the traditional bibliographic database search, such as snowballing (forwards and backwards citation chasing) and communication with relevant topic experts should not be underestimated. Finally, development of a risk-of-bias assessment tool by a panel of methodologists and analysts would be of substantial benefit.

<!-- TODO Ask Matt Page whether he has done any work on this? Studies looking at many risk factors don't report null results in abstract and so do no -->

Talk about problem with studies sharing underlying datasets in two sample Mendelian randomisation frameworks - c.f. EHR databases, which contain the same underlying sample but use different sub-samples on account of the distinct codes/conditions and timepoints used to defined the study cohorts, Mendelian randomisation analysis can be multiple studies using the exact same summary statistics from the same cohorts

One item of particular interest is the attenuation of any effects observed by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the Apoe4 gene region. As covered in the introduction (see Section \@ref()), increasing number of ApoE4 alleles is a major independent risk factor for Alzheimer's disease, and so violates the exclusion restriction criteria of .

Introduce assumptions underlying Mendelian randomisation (with image). Higlight that most if not all included analysis initially describing a protective effect of LDL-c, which attenuates to the NULL once ApoE4 gene regions are removed. Note the region must be quite wide - talk about controversy surrounding Benn paper

Additionally many of the studies suggested a link between, but then in the sensitivity analyses or in the Discussion disclaim that when adjusted for ApoE4, any association was attenuated to the null. This point is important given one of the core assumptions of

MR are limited by the need for SNP to be present in both datasets (for TSMR).

<!-- TODO Talk about discrepancies between results of different methods, but highlight that it is a problem as the information is being contributed by the same people. -->

Recommend future large scale GWAS of other dementia outcomes, notably vascular dementia, or the use of existing GWAS to  <!-- TODO At least one vascular dementia GWAS exists - find it and cite it here. -->

A key example of the importance of . In almost all cases, Mendelian randomisation studies examine. A clear exampe of this is Ben et al, 2017, where the ApoE variants were not sufficient identified and excluded, and the published paper detailed evidence for a protective effect of LDL-c was identified (`estimate(0.83, .75,.92,"RR")`).[@benn2017] Following several rapid responses, the data was re-analysed excluding a larger area around ApoE4 which attenuated this finding towards the null.[@benn2017a]

@@

As highlighted in Section \@ref(diverse-sources-preprints), this review explicitly sought to synthesize evidence across different publication statuses (preprinted vs. published). Using the tool described in Chapter \@ref(sys-rev-tools-intro), two preprint serves related to health and biomedical sciences were search as part of this review.  The small number of studies return by the searches (or the absence of any hits in the medRxiv database - see Figure \@ref(fig:prisma-flow-fig)) is due to the timing of the preprint searches. The searches for this review were performed in mid-July 2019, but the medRxiv repository, an offshoot of the Epidemiology and Clinical Trials categories of the bioRxiv preprint server, only registered its first preprint 25th June 2019. As such, at the point it was searched, the medRxiv database contained only a very small number of records (n=`r n_at_search["med"]`).

@@

```{r preprintGrowth, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:preprintGrowth-cap)', out.width='100%', fig.scap='(ref:preprintGrowth-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/preprint_growth.png"))
```
<!----------------------------------------------------------------------------->

Three relevant preprints from the bioRxiv hits were identified. The added evidential value of including these preprints was described in Section \ref(sys-rev-including-preprints-res), and indicated that results available only via preprinted reports can contribute substantially to a meta-analysis. Of note, all three included preprints described Mendelian randomisation analysis, potentially indicating that more biologically-focused study designs are over-represented in the bioRxiv repository.

Of the three identified preprints, two were subsequently published as of September 2020. This fits well with the analysis presented in Section \@ref(medrxivr-limitations) that, allowing for a two-year lag, approximately two-thirds of preprints are published, and nicely illustrates the dual advantages of preprinted reports to evidence syntheses. Firstly, preprints provide an advance snapshot of the literature, capturing preprinted articles that are subsequently published but were not . Secondly, inclusion of preprints allows for results that may never be formally published to be included in an evidence synthesis exercise.

Consider the example of one of the Mendelian randomisation analyses of the effect of LDl-c on Alzheimer's disease, which found no impact of LDL-c on AD following removal of APOE4. The manuscript was initially published in bioRxiv in July 2017[@zhu2017] which was subsequently published in Nature Communications in January 2018, following peer-review.[@zhu2018] While this study was captured by both the preprint and published searches in this review, had the searches been run within this window, the preprint would have contributed unique data to the review. This illustrates that while it may not have aided this review, if the aim is to find the current state of the art in the topic area at the time of searching, inclusion of preprints is a necessity.

More recently, inclusion of preprints in systematic reviews has become significantly more widespread. This is largely due to the role of preprint servers, in particular medRxiv, as a key evidence dissemination venue during the early stages of the COVID-19 pandemic. How well this adoption of preprints will transfer to other less-urgent topics, where the speed of research does not put the same focus on preprinted articles, is currently unknown.

>>>>

# Systematic review of all available evidence on the association between blood lipids and dementia outcomes {#sys-rev-heading}

@@

A straight-forward random-effects meta-analysis model was employed to combine the identified results. A random-effects model was chosen due to the ...

<!-- TODO Finish this description -->

Results were stratified into subgroups on the basis of the overall risk of bias 
assessment, and summary estimates for each subgroup, in addition to an overall effect estimate, are displayed. Additional descriptive statistics are presented, while prediction intervals are shown as a dotted line around the overall effect estimate. Finally, a test subgroup differences between studies at different levels of risk of bias was performed (see subsequent Section \@ref(sys-rev-visualising-results)). All models were implemented using the `metafor` R package.

@@

@@

Following best practice, studies judged to be at critical risk of bias should be excluded from any quantitative analyses. <!-- TODO CITATION NEEDED --> `r num_to_text(n_critical,T)` observational studies were excluded on this basis, predominantly due to a lack of adjustment for any confounders (i.e. the study reported unadjusted estimates).`r critical_citations`

@@

`r num_to_text(obsHyperchol$VaD$n,T)` studies reported on the association of hypercholesterolemia with vascular dementia and provided weak evidence for an effect (`r obsHyperchol$VaD$estimate`, Figure \@ref(fig:obsHyperVaD))

@@

@@

```{r doseResponse, include = FALSE}
```

There were `r n_dr_initial` studies that were eligible for dose-response meta-analysis as they provide several categories of lipid exposure. However, following data extract, `r num_to_text(n_dr_excluded)` were excluded as they did not reported the relevant information needed (most commonly, the cut-off measures for each category).

Across the remaining `r num_to_text(n_dr_final)` studies, a sufficient number of studies (n $\geqslant$ 3) were identified only for the total cholesterol-Alzheimer's, LDL-Alzheimer's, and total cholesterol-dementia strata. This analysis provided very weak evidence for a non-linear effect of lipid levels on dementia outcomes, and Figure \@ref(fig:lipidsDoseResponse) illustrates this for the total cholesterol-Alzheimer's strata. Similar figures for the other analysed lipid-outcome strata are presented in Appendix \@ref(hold).

<!----------------------------------------------------------------------------->
(ref:lipidsDoseResponse-cap) Dose-response meta-analysis of total cholesterol on Alzheimer's disease

(ref:lipidsDoseResponse-scap) Dose-response meta-analysis of total cholesterol

```{r lipidsDoseResponse, echo = FALSE, results="asis", fig.pos = "H",fig.align='center', fig.cap='(ref:lipidsDoseResponse-cap)', out.width='70%', fig.scap='(ref:lipidsDoseResponse-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/dr_AD_TC.png"))
```
<!----------------------------------------------------------------------------->

@@

Data on another potentially important source of heterogeneity.

Plus other such as education level and baseline cognitive scores, but data were either not reported for most studies or when reported, were too diverse to synthesize.

@@

@@

The risk of bias due to missing evidence in each synthesis is shown beside the overall summary diamond in each forest plot presented above. For randomised controlled trials and non-randomised studies of interventions, the risk of bias due to missing evidence was assessed to be minimal. However, there was substantial evidence that results were selectively reported in studies examining the effect of lipid fractions on dementia outcomes (Figures \@ref(fig:lipidFractionsDementia), \@ref(fig:lipidFractionsAD) and \@ref(fig:lipidFractionsVaD).

@@

As show in Figure \@ref(fig:prisma-flow-fig), the number of hits returned by the preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). From these hits, three preprinted reports of eligible studies were included in the review, of which two described unique studies not captured by the main search.[@andrews2019; @so2017]

Including preprints did provide useful additional evidential value in a number of meta-analysis. To demonstrate this, the effect of HMGCR SNPS on Alzheimer's disease (Figure \@ref(fig:mrStatinADFig)) in Mendelian randomisation studies was re-analysed using a fixed-effect model. Examination of the weight assigned to each result in the analysis illustrates that a large proportion (`r mrStatinPreprintWeight`) of the weight is given to the preprinted result.

@@

<!-- <!-----------------------------------------------------------------------------> -->
<!-- (ref:preprintWeights-caption) Amount of information added by inclusion of preprinted results, estimated from the weight assigned to the study in a random-effect meta-analysis  -->

<!-- (ref:preprintWeights-scaption) Added information due to inclusion of preprints -->

<!-- ```{r preprintWeights-table, message=FALSE, results="asis", echo = FALSE} -->
<!-- ``` -->
<!-- <!-----------------------------------------------------------------------------> -->

@@

Investigation of the publication status of the two preprints reporting studies not also identified by the main search after a two-year lag found that one had been subsequently published in late 2019.[@andrews2019; @andrews2021] The final preprint has not yet been published.[@so2017]

@@

<!-- TODO Check inclusion of Rockwood 2002:  -->
This review has presented a summary of the available evidence on the association between lipids, and treatments that modify lipids such as statins, and the subsequent risk of dementia. This discussion seeks to  to summarise the key findings in terms of literature sources and results as reported. A detailed comparison across the evidence sources, exposure measures and sources of bias reported here is presented as part of the triangulation exercise in Chapter \@ref(tri-heading).

@@

There was some evidence of protective effect of statins on all-cause and Alzheimer's disease dementia when looking at solely at observational studies. This finding was not supported by evidence from the two available RCTs, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings may be a result of heterogeneity in exposure (e.g. mid-life in studies of lipids with late-life lipid reduction in RCTs) or alternatively due to biases within the non-randomised studies.

<!-- Some evidence that age has impact on observed lipid-dementia relationship - use to link  -->

The majority of studies were non-randomised studies of lipids, or treatments that affect lipid levels such as statins. This distribution of evidence between analytical designs is to be expected. Randomised controlled trials of dementia are particularly challenging, as the long follow-up made necessary by the long latent period of the condition, makes trials logistically challenging and financial expensive. Similarly, Mendelian randomisation is a comparatively new study design (as illustrated in Figure \@ref(fig:typeByYear)), which appears in the literature in recent years, driven by the availability of summary genome wide association studies (GWAS) that form the basis of the two-sample Mendelian randomisation approach.

@@

A common theme was an absence of studies examining vascular dementia as an outcome, most noticeable when comparing the evidence base for statins in dementia/Alzheimer's (Figures \@ref(fig:obsStatinDementiaFig) & \@ref(fig:obsStatinADFig)) with that available for vascular dementia (Figure \@ref(fig:obsStatinVaDFig)). This is particularly interesting given that lipids and statins are strongly related to the prevention of vascular disease. A potential explanation for this observation may be publication bias or the "file-drawer effect",[@rosenthal1979] though there was very weak evidence of a small-study effects for this outcome (of which publication bias is a potential cause).[@sterne2011] Similarly, only one Mendelian randomisation studies examined this outcome, primarily because of the absence (until recently) of vascular dementia GWAS which precludes a two-sample approach.

Of note, this review did not include the commonly cited PROSPER RCT, which examined the effect of pravastatin on CVD risk, reporting on cognitive outcomes as one of several secondary outcomes.[@shepherd2002a] While widely cited in relation to the effect of statins on dementia risk and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial reported solely on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up. Though an useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using recognised criteria as cognitive tests should feed into a broader diagnostic pathway (see Section \@ref(diagnostic-criteria)). As such, this trial did not met the inclusion criteria for this review.

@@

Risk of bias across the individual results was generally quite high, and the causes and expected directions of these biases are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of statin levels. In many cases, estimates were know to be missing from meta-analyses not at random due to preferential reporting of significant results observed in a number of analysis, leading to the high-risk judgement. These missing estimates were most commonly identified via analysis of conference abstract/final publication pairs.[@yamada2009; @yamada2009a] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the Table.").[@ancelin2013] However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analysis (which provided very weak evidence for an effect), other than increasing the precision of the summary estimate.

Finally in terms of generalisability, despite a large proportion of included studies being conducted in the Western world (Figure \@ref(fig:)), the applicability of my results to other populations is aided by th inclusion of studies which made use of data from the Taiwan health insurance database.

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018a; @kuzma2018a] However, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include[@chu2018;@power2015] or used a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

I identified one previous review of Mendelian randomisation studies examining risk factors for Alzhiemer's disease. However, this review was conducted prior to the majority of Mendelian randomisation studies included in this review being published and extracted results including SNPs in the APoE4 genetic region (see the following section for a discussion of the bias this introduces).

Despite these differences in time-scales and methodology, the duplication of work across reviews (including this review) is substantial. In retrospect, an alternative approach to conducting a further systematic review from scratch could have been employed. Known as an umbrella review, or review-of-reviews, these studies use other systematic reviews rather than primary studies as the unit of analysis.[@aromataris2015; @smith2011] This approach would have enabled more efficient identification of relevant primary studies to which the methods which sets this review apart from other published reviews could have been applied.

@@

One of the particular strengths of this review is the inclusion and critical assessment of Mendelian randomisation studies as a source of evidence.

Mendelian randomisation is a powerful analytical technique, using natural variation in participants genomes to identify causal links between a genetically determined risk factor and an outcome, given that the core assumptions detailed in Figure \@ref(fig:mrAssumptions) are valid. However, inclusion of Mendelian randomisation as an acceptable study design in this review was complicated by a number of factors.

@@

```{r mrAssumptions, echo = FALSE, results="asis", fig.pos = "H",fig.align='center', fig.cap='(ref:mrAssumptions-cap)', out.width='70%', fig.scap='(ref:mrAssumptions-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/mrAssumptions.png"))
```
<!----------------------------------------------------------------------------->

Firstly, this study design is relatively new, particularly when compared to randomised trials or cohort studies. Figure \@ref(fig:typeByYear) demonstrates that Mendelian randomisation studies only begin to appear in the evidence base much later than NRSE/NRSI, likely due to the limited availabilt of large scale GWAS datasets needed for two-sample Mendelian randomisation analyses. As such, the process and tools for systematically assessing this study design are not as well developed. A key example of this is in the absence of validated search filters for Mendelian randomisations studies. This limitation is further complicated by the varying terminology used to describe the method, particularly in the early years of it's application, which lead o me including general terms for instrumental variable analyses in my search.

Additionally, there is currently no widely used risk-of-bias assessment tool for Mendelian randomisation studies. A recent commentary provided a checklist  interpreting Mendelian randomisation studies, this guide includes reporting items in their quality checklist. While reporting quality is important, it is a separate consideration to internal validity, as discussed in Section \@ref(risk-of-bias). Similarly, a previous review of Mendelian randomisation studies used the Q-Genie tool which was validated to assess the quality of genetic association studies in meta-analysis.[@sohani2015] While this tool addresses the studies used, it does not assess the additional methodological considerations of the Mendelian randomisation analysis itself. <!-- TODO Read this tool. --> For this review, I utilised the best available author-devised tool, sourced from a recent review of systematic reviews of Mendelian randomisation studies.

@@

Useful future work to improve the methodology for inclusion of Mendelian randomisation studies in systematic reviews should involve the development of a validated search filter for this study design.[@waffenschmidt2020; @wagner2020] Alternatively, in better-resourced reviews, a dedicated search for "risk factors" and "dementia" and "Mendelian randomisation" followed by manual review of studies that look across multiple risk factors would be advisable. This was not feasible in the context of this review, given the large number of records to be screened even when using study design filters (n=`r comma(16109)`). Additionally, the value of methods that supporting the traditional bibliographic database search, such as snowballing (forwards and backwards citation chasing) and communication with relevant topic experts should not be underestimated. Finally, development of a risk-of-bias assessment tool for this study design by a panel of methodologists and analysts would be of substantial benefit to the field.

One item of particular interest is the attenuation of any effects observed by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the Apoe4 gene region. As covered in the introduction (see Section \@ref(intro-basic-science)), increasing number of ApoE4 alleles is a major independent risk factor for Alzheimer's disease, and so violates the exclusion restriction criteria (Figure \@ref(fig:mrAssumptions)). In all cases, excluding these variants attenuates the observed effect to the null. A clear example of this is Benn _et al_ (2017), where the ApoE variants were not sufficient identified and excluded and the published paper detailed evidence for a protective effect of LDL-c was identified (`estimate(0.83, .75,.92,"RR")`).[@benn2017] Following several rapid responses, the data was re-analysed excluding a larger area around ApoE4 which attenuated the finding to the null.[@benn2017a]

@@

As highlighted in Section \@ref(diverse-sources-preprints), this review explicitly sought to synthesize evidence across different publication statuses (preprinted vs. published). Using the tool described in Chapter \@ref(sys-rev-tools-intro), two preprint serves related to health and biomedical sciences were search as part of this review.  The small number of studies returned by the searches (or the absence of any relevant hits in the medRxiv database - see Figure \@ref(fig:prisma-flow-fig)) is due to the timing of the preprint searches. The searches for this review were performed in mid-July 2019, but the medRxiv repository, an offshoot of the Epidemiology and Clinical Trials categories of the bioRxiv preprint server, only registered its first preprint 25th June 2019. As such, at the point it was searched, the medRxiv database contained only a very small number of records (n=`r n_at_search["med"]`).

@@

```{r preprintGrowth, echo = FALSE, results="asis", fig.pos = "H",fig.align='center',fig.cap='(ref:preprintGrowth-cap)', out.width='80%',fig.align='center',fig.scap='(ref:preprintGrowth-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/preprint_growth.png"))
```
<!----------------------------------------------------------------------------->

Three relevant preprints from the bioRxiv hits were identified. The added evidential value of including these preprints was described in Section \@ref(sys-rev-including-preprints-res), and indicated that results available only via preprinted reports can contribute substantially to a meta-analysis. Of note, all three included preprints described Mendelian randomisation studies, potentially indicating that more biologically-focused study designs are over-represented in the bioRxiv repository.

Of the three identified preprints, two were subsequently published as of September 2020. This fits well with the analysis presented in Chapter \@ref(sys-rev-tools-heading) that, allowing for a two-year lag, approximately two-thirds of preprints are published, and also nicely illustrates the dual advantages of preprinted reports to evidence syntheses.

Firstly, preprints provide an advance snapshot of the literature, capturing articles that will eventually be published but were not available at the time of the main search. Consider the example of one eligible preprint in this review was initially posted on bioRxiv in July 2019[@andrews2019b] and was subsequently published in 2021 following peer-review.[@andrews2021] Secondly, inclusion of preprints allows for results that may never be formally published to be included in an evidence synthesis exercise as is the case with the second preprint included in this review.[@so2017] Both of these aspects illustrate that, if the aim is to find the current state of the art in the topic area at the time of searching, inclusion of preprints is a necessity.

More recently, inclusion of preprints in systematic reviews has become significantly more widespread. This is largely due to the role of preprint servers, in particular medRxiv, as a key evidence dissemination venue during the early stages of the COVID-19 pandemic.[@fraser2020preprinting] How well this adoption of preprints will transfer to other less-urgent topics, where the speed of research does not put the same focus on preprinted articles, is currently unknown.
